Predicting computational reproducibility of data analysis pipelines \\in large population studies using collaborative filtering

### 2018-07-24 paper.tex
Added missing files, generated figures from updated data, various corrections until section V

\begin{abstract}
With the rise of the reproducibility crisis in several scientific 
domains, evaluating the computational reproducibility of data analysis 
pipelines has become a critical issue. It is, however, a cumbersome 
process for analyses that involve data from large populations of 
subjects, due to the computational and storage requirements. 
Nevertheless, executions of a given analysis pipeline are not 
independent, as they are obtained from the same program. We leverage 
this observation to propose a method to predict the computational 
reproducibility of data analysis pipelines in large population studies. 
We formulate the problem as a collaborative filtering problem with 
constraints on the construction of the training set. We propose 6 
different strategies to build the training set and evaluate them on 2 
datasets, a synthetic one modeling a population with a growing number 
of subject types, and a real one obtained on neuroinformatics 
pipelines. Results show that one sampling method, ``Random File Numbers 
(Uniform)" is able to predict computational reproducibility in 
pipelines with a good accuracy. We also analyze the relevance of 
including file and subject biases in the collaborative filtering model. 
We conclude that the proposed method is able to speed-up reproducibility 
evaluations substantially, with a reduced accuracy loss.
\end{abstract}

@@

Our goal is to predict the outcome of reproducibility evaluations in a 
large population of subject from a reduced set of pipeline executions. 
More precisely, we aim at predicting whether a particular file produced 
by an analysis pipeline will be identical across execution conditions, or if it will contain reproducibility \emph{errors}. 
We approach the problem from the point of view of collaborative 
filtering, inspired by previous examples that applied this method 
outside of its initial application domain, for instance 
~\cite{feng2013efficient}.

The paper makes the following contributions:
\begin{enumerate}
\item We model reproducibility evaluations in data processing pipelines as a collaborative filtering problem.
\item We propose strategies to sample the training set under time constraints.
\item We evaluate and compare our sampling strategies on synthetic and real datasets.
\end{enumerate}
The paper ends with a discussion about \todo{x} and a conclusion.

\section{Method}

\subsection{Problem formulation}

The pipeline to be evaluated is represented by a matrix $U$ of size $N_f \times N_s$
where columns represent data coming from different subjects and rows represent 
the files produced by the pipeline execution. Subjects are unordered, 
i.e., columns may be reordered arbitrarily, but files are ordered, for 
instance from their last modification time in a sequential execution, 
and we assume that this order is consistent across subjects. $U_{i, j}$ 
represents the 
reproducibility of file $i$ produced during the processing of subject 
$j$ in two conditions, for instance two different operating systems. In our 
experiments, we restrict $U_{i,j}$ to be boolean, but the same prediction method
can be applied for real values.

Our goal is to predict the test set $\mathbb{T'}$ of missing values of 
$U$ from a training subset $\mathbb{T}$ of known ones. In contrast with the traditional collaborative filtering 
problem, (1) we have control over the construction of the training set, and (2)
the construction of the training set is constrained by the order on the matrix rows, which is formalized as follows: 
\begin{equation}
\begin{array}{l}
\forall (i, j, k) \in \llbracket 1, N_f \rrbracket \times \llbracket 1, N_f \rrbracket \times \llbracket 1, N_s \rrbracket, \\
 \left( U_{i,k} \in \mathbb{T} \quad \mathrm{and} \quad U_{j,k} \in \mathbb{T'} \right) \Rightarrow i < j. \label{eq:time}
 \end{array}
\end{equation}

@@

Collaborative filtering is a technique to predict unknown values of a 
matrix called "utility matrix" from the known ones. Traditionally, the 
matrix represents the ratings of items, represented in columns, by 
users, represented in rows. Ratings might be explicit, when users 
provide ratings through a dedicated system, or implicit, when users' 
behaviors are tracked and analyzed to estimate their preferences. An 
overview of collaborative filtering is provided 
in~\cite{leskovec2014mining}. In our context, the utility matrix is the
pipeline matrix described previously.

@@

A third class of methods, which is the one we will apply to our 
problem, is based on the factorization of the utility matrix to 
estimate latent factors along which users and items are represented. 
This method is described 
in~\cite{koren2009matrix} and became famous as it contributed to winning the 
Netflix prize in 2009. To summarize, the method aims at finding $q_i$ 
and $p_u$ vectors of $\mathbb{R}^f$, where $f$ is the number of latent factors, such that:
\begin{equation*}
r_{ui} = q_i^Tp_u,
\end{equation*}
where $r_{ui}$ is the rating of item $i$ by user $u$. In practice, the optimization
involves a regularization term to avoid overfiting particular users or items. The method
finds $q_i$ and $p_u$ that minimize the following objective:
\begin{equation*}
\sum_{(i,u) \in \mathbb{T}}\left( r_{ui} - q_i^Tp_u\right)^2+\lambda \left( \|{q_i}\|^2 + \|{p_u}\|^2\right)
,
\end{equation*}
where $\lambda$ is the regularization parameter and $\mathbb{T}$ is the training set.

@@

Two good techniques to perform the optimization are stochastic gradient descent and alternating least squares (ALS). 
In the remainder of this paper, we use matrix factorization as implemented in 
Apache Spark's MLLib version 2.3.0, which uses ALS.

% Explain our particular context.
% What our matrix represents
% Explain that you have binary classes (rounding)
% Explain the sampling issue.

\subsection{Sampling of the training set}
\label{sec:training}
As explained before, the training set in our problem cannot be 
constructed by unconstrained random sampling of the matrix.
To address this issue, we investigated the following sampling techniques.

% 2. columns
\paragraph{Complete Columns (Figure~\ref{fig:Columns-Sample-Training-set})}
The training set is sampled by randomly selecting complete columns in
the matrix, that is, complete subject executions. The last selected
column might be incomplete to meet the exact training ratio. This 
method is realistic: it corresponds to a situation where the 
collaborative filtering method will predict the reproducibility in the 
remaining subjects from the subjects in the training set.

% 3. rows
\paragraph{Complete Rows (Figure~\ref{fig:Rows-Sample-Training-set})} The training set is sampled by selecting complete 
rows in the matrix, that is, the first files produced by
every execution. The last selected row might be incomplete to meet the
exact training ratio. This method is realistic: it corresponds to a 
situation where the processing of all the subjects is launched and 
interrupted before the execution is complete. The collaborative 
filtering method will then predict the reproducibility of the remaining 
files.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_columns_04_training}
                  \caption{Complete Columns}
                  \label{fig:Columns-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_rows_04_training}
                  \caption{Complete Rows}
                  \label{fig:Rows-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                 \includegraphics[width=\columnwidth]{figures/5vs7_random-real_04_training}
                \caption{Random Subjects (RS)}
                  \label{fig:Uniform-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_diagonal_04_training}
                  \caption{RFNU}
                  \label{fig:Diagonal-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_random-triangular-largest_04_training}
                  \caption{RFNT-L}
                  \label{fig:triangular-L-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_random-triangular-smallest_04_training}
                  \caption{RFNT-S}
                  \label{fig:triangular-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
        \includegraphics[width=\columnwidth]{figures/5vs7_random-unreal_04_training}
                  \caption{Random Unreal}
                  \label{fig:Random-Unreal-Sample-Training-set}
        \end{subfigure}
        \caption{Sample training set for different methods with a training ratio of 0.4.}                
\end{figure}

\paragraph{Random Subjects -- RS (Figure~\ref{fig:Uniform-S-Sample-Training-set})} This method builds the 
training set by selecting the files from random subjects 
until the training ratio is reached. The file selected in a 
subject is the file with the lowest index in this subject that has not 
been already selected in the training set. This method is realistic as 
files are sampled according to their production time-stamps.

@@

          \textit{a} = (2\alpha - 1) N_{f} & \text{if $\alpha > 0.5$}
  \end{cases}
\]
 For $\alpha > 0.5$, this ensures that the average number of files in 
 the training set is $\alpha N_f$. For $\alpha \leq 0.5$, the number of 
 files selected in every subject is sampled from $U(0,N_f)$ with 
 probability $2\alpha$, and set to 0 otherwise.
RFNU is realistic as files are sampled according to their production time-stamps.
Due to sampling issues, it is possible that the actual training ratio obtained with this method
does not match the target one. We check that the difference between the target and real training ratios
was lower than 0.01.\\

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution $T(a, b, c)$ as in Figure~\ref{fig:triangular}. The mean of the distribution is 
$\frac{a+b+c}{3}$. We set \textit{c} to $N_{f}$ and we set 
\textit{a} and \textit{b} with two approaches that ensure that the 
average of the distribution is $\alpha N_{f}$ ($\alpha$ is the training 
ratio):
\begin{enumerate}
        \item Largest a (RFNT-L, Figure~\ref{fig:triangular-L-Sample-Training-set}): a is set to the 
        largest possible value, i.e., b. The average of the 
        distribution is $\frac{2a+N_{f}}{3}$, therefore 
        $a=\frac{3\alpha-1}{2}N_{f}$. When $\alpha \leq 1/3$, the 
        number of files in every subject is sampled from $T(0, 0, N_f)$ 
        with probability $3\alpha$ and set to 0 otherwise, to ensure that the average number of files
        in the training set is $\alpha N_f$.
        \item Smallest a (RFNT-S, Figure~\ref{fig:triangular-S-Sample-Training-set}): a is 
        set to the smallest possible value, i.e., 0. The average of the 
        distribution is $\frac{b+N_{f}}{3}$, therefore 
        $b=N_{f}(3\alpha-1)$. This implies that $\alpha \geq 1/3$, so 
        that $b \geq 0$, and $\alpha \leq 2/3$, so that $b \leq N_f$. 
        When $\alpha < 1/3$, we set $b=0$ and the number of files 
        selected in every subject is sampled from $T(0, 0, N_f)$ with 
        probability $3\alpha$ and set to 0 otherwise, as in RFNT-L. 
        When $\alpha \geq 2/3$, we set $b=N_f$ and the number of files 
        selected in every subject is sampled from $T(0, N_f, N_f)$ with 
        probability $3(1-\alpha)$ and set to $N_f$ otherwise, still to 
        ensure that the average number of files in the training set is 
        $\alpha N_f$.
\end{enumerate}
As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the motivation for the 
RFNT-L method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set, which is not the case for RFNU.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

% 0. random unreal (baseline)
\paragraph{Random Unreal (Figure ~\ref{fig:Random-Unreal-Sample-Training-set})}
The training set is sampled in a random uniform way, 
regardless of the file generation times. This method will be used as a
baseline for comparison with other methods, although it is not
realistic.

@@

\section{Dataset}

@@

We generated synthetic matrices as shown in 
Figure~\ref{fig:synthetic-data}. Each matrix has 100 files and 100 
subjects of different \emph{types}. Subjects of the same type behave 
identically and all types contain the same number of subjects $\pm$ 1 
\todo{Soudabeh: check that}. Such a decomposition by subject type 
correspond to the case where different subjects may have data of 
different nature, as it is for instance the case in the neuroimaging 
dataset of the Human Connectome Project~\cite{van2013wu} where not all the 
subjects have the same amount of images.

For subjects of a given type, matrix elements vary in $log(n)$  blocks,         
where $n$ is the number of types, with all the possible variation 
patterns: some types do not vary at all, while other ones vary in every 
block. Such variation patterns are meant to mimic the logic of data 
processing pipelines: each block of files represents the files produced 
at a given stage of the pipeline, which may or may not contain 
reproducibility errors depending on the subject type.

\begin{figure}
\centering
        %\begin{subfigure}[b]{0.4\linewidth}
                  %\includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/1_SubjectType_utility_matrix.png}
                  %\caption{1 type}
        %\end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/2_SubjectType_utility_matrix.png}
                  \caption{2 types}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/4_SubjectType_utility_matrix.png}
                  \caption{4 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/8_SubjectType_utility_matrix.png}
                  \caption{8 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/16_SubjectType_utility_matrix.png}
                  \caption{16 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/32_SubjectType_utility_matrix.png}
                  \caption{32 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/64_SubjectType_utility_matrix.png}
                  \caption{64 types}
        \end{subfigure}
\caption{Synthetic Data}
\label{fig:synthetic-data}
\end{figure}

@@

We collected data to evaluate the computational reproducibility of analysis
pipelines of the Human Connectome Project~\cite{glasser2013minimal}. We
processed a set S of 94 subjects randomly selected in the S500 HCP
release~\todo{URL} in three execution conditions with different
versions of the CentOS operating system (5.?, 6.? and 7.?), using the
PreFreesurfer and Freesurfer pipelines
described in~\cite{glasser2013minimal} and available at \todo{URL}. For
each pipeline, we identified the set F of files produced for all
subjects in all conditions. For each condition pair and each pipeline,
we computed a binary difference matrix D of size $|F|\times|S|$, where $D_{i,j}$ is true
if file $i$ of subject $j$ was different in each condition. Rows of
$M$ are ordered by ascending file modification time in the pipeline.

@@

Three experiments have been conducted for each dataset to evaluate the performance of our 
predictions using (1) ALS without biases, (2) ALS with biases, (3) Biases only. 
Results will be evaluated using accuracy, sensitivity and specificity defined as follows:
\[
        Accuracy = \frac{TP+TN}{TP+TN+FP+FN}
\]
\[
        Sensitivity = \frac{TP}{TP+FN}
\]
\[
        Specificity = \frac{TN}{TN+FP}
\]
Where TP is the number of True Positives (correctly predicted reproducibility errors), 
FP is the number of False Positives, TN is the number of true negatives 
and FN is the number of false negatives. \todo{Check if we want to report sensitivity and specificity, otherwise remove.}

We compare our sampling methods to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. \todo{Report model parameters, including number of factors.}

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is 0.5 for all subject types. Random 
Unreal performs very well for all subject types, which confirms that 
ALS is working correctly. Unsurprisingly, all the other methods perform 
better than the dummy classifier, and their accuracy decreases as the 
number of subject types increases.

@@

Results of ALS with bias are reported in Figure~\ref{fig:results-synthetic-als-bias}.
RS, RFNU and RFNTL remain the methods that best compare to Random Unreal, but the achieved
accuracy is much lower than without Biases. This is due to the fact that biases \ldots \todo{finish that.}

>>>>

\begin{abstract}
With the rise of the reproducibility crisis in several scientific 
domains, evaluating the computational reproducibility of data analysis 
pipelines has become a critical issue. It is, however, a cumbersome 
process for analyses that involve data from large populations of 
subjects, due to their computational and storage requirements. 
We present a method to predict the computational 
reproducibility of data analysis pipelines in large population studies. 
We formulate the problem as collaborative filtering, with 
constraints on the construction of the training set. We propose 6 
different strategies to build the training set, and we evaluate them on 2 
datasets, a synthetic one modeling a population with a growing number 
of subject types, and a real one obtained with neuroinformatics 
pipelines. Results show that one sampling method, ``Random File Numbers 
(Uniform)" is able to predict computational reproducibility in 
pipelines with a good accuracy. We also analyze the relevance of 
including file and subject biases in the collaborative filtering model. 
We conclude that the proposed method is able to speed-up reproducibility 
evaluations substantially, with a reduced accuracy loss.
\end{abstract}

@@

Our goal is to predict the outcome of reproducibility evaluations in a 
large population of subjects, from a reduced set of pipeline 
executions. More precisely, we aim at predicting whether a particular 
file produced by an analysis pipeline will be identical across 
execution conditions, or if it will contain reproducibility 
\emph{errors}. We approach the problem from the point of view of 
collaborative filtering, inspired by previous examples that applied 
this method outside of its initial application domain, for instance 
~\cite{feng2013efficient}. The main issue, and originality, of our 
problem lies in the fact that the training set cannot be arbitrarily 
sampled from the utility matrix. Instead, the training set has to 
respect the time constraints that stem from the order in which files 
are produced during pipeline executions.

This paper makes the following contributions:
\begin{enumerate}
\item We model reproducibility evaluations in data processing pipelines as a collaborative filtering problem.
\item We propose strategies to sample the training set under time constraints.
\item We evaluate and compare our sampling strategies on synthetic and real datasets.
\end{enumerate}
The problem formulation, collaborative filtering technique used, and 
proposed sampling strategies for the training set are presented in 
Section~\ref{sec:methods}. The datasets are described in 
Section~\ref{sec:datasets}, and experimental results are reported in 
Section~\ref{sec:results}.
The paper ends with a discussion about \todo{x} and a conclusion.

\todo{Complete the introduction and improve the writing}

\newpage

\section{Method}
\label{sec:methods}
\subsection{Problem formulation}

The pipeline to be evaluated is represented by a matrix $U$ of size 
$N_f \times N_s$ where the $N_s$ columns represent data coming from 
different subjects and the $N_f$ rows represent the files produced by 
the pipeline execution. While subjects are not ordered, files are, for 
instance from their last modification time in a sequential execution. 
We assume that the order on files is consistent across subjects. $U_{i, 
j}$ measures the reproducibility of file $i$ produced during the 
processing of subject $j$ in two conditions, for instance two different 
operating systems. In our experiments, we restrict $U_{i,j}$ to be 
boolean, but our methods can be applied for real values as well.

Our goal is to predict the test set $\mathbb{T'}$ of missing values of 
$U$ from a training set $\mathbb{T}$ of known ones, where $\mathbb{T} 
\cap \mathbb{T'} = \emptyset$ and $\mathbb{T} \cup \mathbb{T'} = 
\{U_{i,j}\}$. In contrast with the traditional collaborative filtering 
problem, (1) we have control over the construction of the training set, 
for instance we can choose to include only files producde by specific 
subjects, or the first files produced by the processing of every 
subject, and (2) the construction of the training set is constrained by 
the 
order on the matrix rows, which is formalized as follows: 
\begin{equation}
\begin{array}{l}
\forall (i, j, k) \in \llbracket 1, N_f \rrbracket \times \llbracket 1, N_f \rrbracket \times \llbracket 1, N_s \rrbracket, \\
 \left( U_{i,k} \in \mathbb{T} \quad \mathrm{and} \quad U_{j,k} \in \mathbb{T'} \right) \Rightarrow i < j. \label{eq:time}
 \end{array}
\end{equation}

@@

Collaborative filtering is a technique to predict unknown values of a 
matrix called "utility matrix" from the known ones. Traditionally, the 
matrix represents the ratings of items, represented in columns, by 
users, represented in rows. Ratings might be explicit, when users 
provide ratings through a dedicated system, or implicit, when users' 
behaviors are tracked and analyzed to estimate their preferences. An 
overview of collaborative filtering is provided 
in~\cite{leskovec2014mining}. In our context, the utility matrix is the
pipeline matrix $U$ described previously.

@@

A third class of methods, which is the one that we will apply to our 
problem, is based on the factorization of the utility matrix to 
estimate latent factors along which users and items are represented. 
This method is described 
in~\cite{koren2009matrix} and became famous as it contributed to winning the 
Netflix prize in 2009. To summarize, the method aims at finding $q_i$ 
and $p_u$ vectors of $\mathbb{R}^f$, where $f$ is the number of latent factors, such that:
\begin{equation*}
r_{ui} = q_i^Tp_u,
\end{equation*}
where $r_{ui}$ is the rating of item $i$ by user $u$. In practice, the optimization
involves a regularization term to avoid overfiting particular users or items. The method
finds $q_i$ and $p_u$ that minimize the following objective:
\begin{equation*}
\sum_{(i,u) \in \mathbb{T}}\left( r_{ui} - q_i^Tp_u\right)^2+\lambda \left( \|{q_i}\|^2 + \|{p_u}\|^2\right)
,
\end{equation*}
where $\lambda$ is the regularization parameter and $\mathbb{T}$ is the training set.

@@

Two good techniques to perform the optimization are stochastic gradient 
descent and alternating least squares (ALS). In the remainder of this 
paper, we use matrix factorization as implemented in Apache Spark's 
MLLib version 2.3.0, which uses ALS. We round predictions to the 
nearest integer to obtain binary values.

\subsection{Sampling of the training set}

\label{sec:training}

As explained before, the training set in our problem cannot be 
constructed by unconstrained random sampling of the matrix. Instead, 
the training and test sets have to comply to Equation~\ref{eq:time}. To 
address this issue, we investigated the following sampling techniques, illustrated in
Figure~\ref{fig:sampling}.

\paragraph{Complete Columns 
(Figure~\ref{fig:Columns-Sample-Training-set})} The training set is 
sampled by randomly selecting complete columns in the utility matrix, 
that is, complete subject executions. The last selected column might be 
incomplete to meet the exact training ratio. This method respects the 
time constraints. It corresponds to a situation where the collaborative 
filtering method will predict the reproducibility of the subjects in 
the test set from the subjects in the training set.

\paragraph{Complete Rows (Figure~\ref{fig:Rows-Sample-Training-set})} 
The training set is sampled by selecting complete rows in the utility 
matrix, that is, the first files produced by every execution. The last 
selected row might be incomplete to meet the exact training ratio. This 
method respects the time constraints. It corresponds to a situation 
where the processing of all the subjects is launched and interrupted 
before the execution is complete. The collaborative filtering method 
will then predict the reproducibility of the remaining files.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_columns_04_training}
                  \caption{Complete Columns}
                  \label{fig:Columns-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_rows_04_training}
                  \caption{Complete Rows}
                  \label{fig:Rows-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                 \includegraphics[width=\columnwidth]{figures/5vs7_random-real_04_training}
                \caption{Random Subjects (RS)}
                  \label{fig:Uniform-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_diagonal_04_training}
                  \caption{RFNU}
                  \label{fig:Diagonal-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_random-triangular-largest_04_training}
                  \caption{RFNTL}
                  \label{fig:triangular-L-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{figures/5vs7_random-triangular-smallest_04_training}
                  \caption{RFNTS}
                  \label{fig:triangular-S-Sample-Training-set}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
        \includegraphics[width=\columnwidth]{figures/5vs7_random-unreal_04_training}
                  \caption{Random Unreal}
                  \label{fig:Random-Unreal-Sample-Training-set}
        \end{subfigure}
        \caption{Sample training set for different methods with a training ratio of 0.4.}
        \label{fig:sampling}
\end{figure}

\paragraph{Random Subjects -- RS 
(Figure~\ref{fig:Uniform-S-Sample-Training-set})} This method builds 
the training set by selecting the files from random subjects until the 
training ratio is reached. The file selected in a subject is the file 
with the lowest index in this subject that has not been already 
selected in the training set. This respects the time constraints as files are 
sampled according to their production time-stamps.

@@

          \textit{a} = (2\alpha - 1) N_{f} & \text{if $\alpha > 0.5$}
  \end{cases}
\]
 For $\alpha \leq 0.5$, we ensure that the average number of selected 
 files by subject is $\alpha N_f$ by sampling the number of 
 files selected in every subject from $U(0,N_f)$ with 
 probability $2\alpha$, and setting it to 0 otherwise. For $\alpha > 
 0.5$, this is ensured by the value used for $a$, which leads the 
 average of U(a, b) to be $\alpha N_f$.

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution $T(a, b, c)$ as in Figure~\ref{fig:triangular}. The mean of the distribution is 
$\frac{a+b+c}{3}$. We set \textit{c} to $N_{f}$ and we set 
\textit{a} and \textit{b} with two approaches that ensure that the 
average of the distribution is $\alpha N_{f}$:
\begin{enumerate}
        \item Largest a (RFNTL, Figure~\ref{fig:triangular-L-Sample-Training-set}): a is set to the 
        largest possible value, i.e., b. The average of the 
        distribution is $\frac{2a+N_{f}}{3}$, therefore 
        $a=\frac{3\alpha-1}{2}N_{f}$. When $\alpha \leq 1/3$, the 
        number of files in every subject is sampled from $T(0, 0, N_f)$ 
        with probability $3\alpha$ and set to 0 otherwise, to ensure 
        that the average number of files by subject in the training set 
        is $\alpha N_f$.
        \item Smallest a (RFNTS, Figure~\ref{fig:triangular-S-Sample-Training-set}): a is 
        set to the smallest possible value, i.e., 0. The average of the 
        distribution is $\frac{b+N_{f}}{3}$, therefore 
        $b=N_{f}(3\alpha-1)$. This implies that $\alpha \geq 1/3$, so 
        that $b \geq 0$, and $\alpha \leq 2/3$, so that $b \leq N_f$. 
        When $\alpha < 1/3$, we set $b=0$ and the number of files 
        selected in every subject is sampled from $T(0, 0, N_f)$ with 
        probability $3\alpha$ and set to 0 otherwise, as in RFNTL. 
        When $\alpha \geq 2/3$, we set $b=N_f$ and the number of files 
        selected in every subject is sampled from $T(0, N_f, N_f)$ with 
        probability $3(1-\alpha)$ and set to $N_f$ otherwise, still to 
        ensure that the average number of files by subject in the training set is 
        $\alpha N_f$.
\end{enumerate}
As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the motivation for the 
RFNTL method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set, which is not the case for RFNU or RFNTS.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

\paragraph{Random Unreal (Figure 
~\ref{fig:Random-Unreal-Sample-Training-set})} The training set is 
sampled in a random uniform way, regardless of the file generation 
times. This method does not respect the time constraints in 
Equation~\ref{eq:time}, it will be used as a baseline for comparison 
with other methods.

@@

\section{Datasets}

\label{sec:datasets}

@@

We generated synthetic matrices as shown in 
Figure~\ref{fig:synthetic-data}. Each matrix has 100 files and 100 
subjects of different \emph{types}. Subjects of the same type behave 
identically and all types contain the same number of subjects $\pm$ 1. 
Such a decomposition by subject type correspond to the case where 
different subjects may have data of different nature, as it is for 
instance the case in the neuroimaging 
dataset of the Human Connectome Project~\cite{van2013wu} where not all the 
subjects have the same amount of images.

 For subjects of a given type, the reproducibility matrix consists of 
 $log(n)$  blocks, where $n$ is the number of types. Blocks are 
 arranged with all the possible variation patterns: some types do not 
 vary at all, while other ones vary in every block. Such variation 
 patterns are meant to mimic the logic of data 
processing pipelines: each block of files represents the files produced 
at a given stage of the pipeline, which may or may not contain 
reproducibility errors depending on the subject type.

\begin{figure}
\centering
        %\begin{subfigure}[b]{0.4\linewidth}
                  %\includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/1_SubjectType_utility_matrix.png}
                  %\caption{1 type}
        %\end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/2_SubjectType_utility_matrix.png}
                  \caption{2 types}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/4_SubjectType_utility_matrix.png}
                  \caption{4 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/8_SubjectType_utility_matrix.png}
                  \caption{8 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/16_SubjectType_utility_matrix.png}
                  \caption{16 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/32_SubjectType_utility_matrix.png}
                  \caption{32 types}
        \end{subfigure}
                \begin{subfigure}[b]{0.4\linewidth}
                  \includegraphics[width=\columnwidth]{data/Utility_Matrix/Synthetic/synthetic_subject_types/64_SubjectType_utility_matrix.png}
                  \caption{64 types}
        \end{subfigure}
\caption{Synthetic reproducibility matrices. White cells denote reproducibility errors.}
\label{fig:synthetic-data}
\end{figure}

@@

We collected data to evaluate the computational reproducibility of analysis
pipelines of the Human Connectome Project~\cite{glasser2013minimal}. We
processed a set S of 94 subjects randomly selected in the S500 HCP
release\footnote{\url{https://db.humanconnectome.org}} in three execution conditions with different
versions of the CentOS operating system (5.\todo{?}, 6.8 and 7.2), using the
PreFreesurfer and Freesurfer pipelines described 
in~\cite{glasser2013minimal} and available on 
GitHub\footnote{\url{https://github.com/Washington-University/Pipelines/releases/tag/v3.19.0}}. 
For each pipeline, we identified the set F of files produced for all
subjects in all conditions. For each condition pair and each pipeline, 
we computed a binary reproducibility matrix U of size $|F|\times|S|$, 
where $U_{i,j}$ is true iif file $i$ of subject $j$ was different in 
each condition. Rows of $U$ were ordered by ascending file modification 
time in a random subject in S.

@@

\label{sec:results}

Two experiments have been conducted for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with biases. 
Results will be evaluated using classification accuracy and compared to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. \todo{Report model parameters, including number of 
factors.} Due to sampling issues, it is possible that the actual 
training ratio obtained with some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true.

%~ , sensitivity and specificity defined as follows:
%~ \[
        %~ Accuracy = \frac{TP+TN}{TP+TN+FP+FN}
%~ \]
%~ \[
        %~ Sensitivity = \frac{TP}{TP+FN}
%~ \]
%~ \[
        %~ Specificity = \frac{TN}{TN+FP}
%~ \]
%~ Where TP is the number of True Positives (correctly predicted reproducibility errors), 
%~ FP is the number of False Positives, TN is the number of true negatives 
%~ and FN is the number of false negatives. \todo{Check if we want to report sensitivity and specificity, otherwise remove.}

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is close to 0.5 for all subject types. Random 
Unreal performs very well for all subject types, which confirms that 
ALS is working correctly. Unsurprisingly, all the other methods perform 
better than the dummy classifier, and their accuracy decreases as the 
number of subject types increases.

@@

Results of ALS with bias are reported in 
Figure~\ref{fig:results-synthetic-als-bias}. RS, RFNU and RFNTL remain 
the methods that best compare to Random Unreal, but the achieved 
accuracy is much lower than without Biases. This is presumably due to 
the fact that file biases are all close to 0.5 in the synthetic data. 
In such a situation, biases are detrimental to the accuracy of the 
collaborative filtering process and shouldn't be included.
